---
title: "ECO395M: Data Mining --- Exercise 01"
subtitle: "ECO395M: Data Mining"
author: "Scott Cohn"
date: "`r Sys.Date()`"
header-includes: 
  - \usepackage{float} 
  - \floatplacement{figure}{H}
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_html: default
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
library(tidyverse)
library(tidymodels)
library(tidyquant)
library(kknn)
library(lubridate)
library(scales)
library(patchwork)
library(Hmisc)
library(kableExtra)
library(extrafont)
library(ggthemes)
library(ggmap)
library(tidygeocoder)

# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, 
                      cache.extra = packageVersion('tufte'),
                      fig.pos= "h"
                      )
options(
  htmltools.dir.version = FALSE, 
  scipen = 999 # turn of scientific notation
        )
```

# Data visualization: Gas Prices

```{r gasprices, message=TRUE, warning=FALSE, cache=TRUE, include=FALSE}
gasprices <- read_csv("../../ECO395M/data/GasPrices.csv")

# geocode the addresses
gasprices <- gasprices %>% 
  tidygeocoder::geocode(street = Address,
          city = 'Austin',
          state = 'TX',
          postalcode = Zipcode,
          method = 'cascade', 
          lat = latitude , long = longitude)
```

```{r thA, echo=TRUE, fig.cap="Price with competition", fig.height=3.33, fig.width=5, cache=TRUE}
# boxplot
gasprices %>% 
  ggplot( aes(x=Competitors, y=Price, fill=Competitors)) + 
    geom_boxplot() + 
    scale_fill_brewer(type="qual") +
    geom_jitter(color="black", size=0.6, alpha=0.9) + 
    theme_minimal() + 
    theme(text=element_text(family="Palatino")) + 
    labs(y="Price") 
```

`r newthought("Theory A:")` Gas stations charge more if they lack direct competition in sight.


`r newthought("Conclusion:")` The cheapest gas stations seem to charge similarly. Gas stations that have no competition have higher prices at the 50th and 75th percentile. There are several gas stations, again facing no competition, pricing 10 cents per gallon higher than all of the rest.

\pagebreak 

```{r thB, echo=TRUE, fig.cap="Price with income", fig.height=4.6, fig.pos='h!', fig.width=6, cache=TRUE}
# scatter plot
gasprices %>% 
  ggplot(aes(x=Income, y=Price, color=Brand)) + 
  geom_jitter() + 
  scale_x_continuous(breaks=seq(10000,130000,15000)) +
  scale_color_brewer(palette="Set1") +
  geom_rangeframe(color="black") +
  theme_tufte()
```

`r newthought("Theory B:")` The richer the area, the higher the gas price.


`r newthought("Conclusion:")` There is a lot of variation within brand across the income spectrum. There does seem to be a trend that higher income areas do have higher prices per gallon at the gas station.

\pagebreak 

```{r thC, echo=TRUE, fig.cap="Shell versus competition", fig.height=3, fig.width=7, cache=TRUE}
# Bar plot
gasprices %>% 
  ggplot( aes(x=Brand, y=Price, fill=Brand) ) +
  stat_summary(fun.data=mean_sdl, geom="bar") +
  stat_summary(fun.data=mean_cl_boot, geom="errorbar", width=0.3) +
  labs(x="") +
  coord_flip() +
  scale_fill_brewer(palette="Set1") +
  theme_tufte()
```

`r newthought("Theory C:")` Shell charges more than other brands.


`r newthought("Conclusion:")` Visually, the averages are very close; Shell seems to price similarly to other brands. I would not conclude, visually, that Shell prices notably higher than the rest. 

\pagebreak 

```{r thD, echo=TRUE, fig.cap="Price with stoplights", fig.fullwidth=TRUE, fig.height=4, fig.width=9, cache=TRUE}
# faceted histogram
gasprices %>% 
  ggplot( aes(x=Price, fill=Stoplight) ) + 
  geom_histogram(position="identity", bins=15) + 
  facet_grid(. ~ Stoplight) + 
  scale_fill_brewer(type="qual") +
  labs(y="Count") + 
  geom_rangeframe() +
  theme_tufte()
```

`r newthought("Theory D:")` Gas stations at stoplights charge more.


`r newthought("Conclusion:")` The distribution of price not at a stoplight is right-skewed whereas the distribution of price at a stoplight appears to be more normally distributed (maybe even left-skewed). Limited data makes this a bit difficult to parse. They seem to be similarly centered around 1.8ish. It seems that gas stations as stoplights price a bit higher than their non-stoplight counterparts.

\pagebreak 

```{r thE, echo=TRUE, fig.cap="Price with highway access", fig.fullwidth=TRUE, fig.height=3, fig.width=9.5, cache=TRUE}
# any
p1 <- gasprices %>% 
  ggplot( aes(x=Price, fill=Highway)) + 
    geom_density(alpha=0.6) + 
    scale_fill_brewer(type="qual") +
    theme_tufte() +
    labs(y="Density")

p2 <- gasprices %>% 
  ggplot( aes(x=Highway, y=Price, fill=Highway)) + 
    geom_boxplot() + 
    scale_fill_brewer(type="qual") +
    geom_jitter(color="black", size=0.4, alpha=0.9) + 
    theme_tufte() + 
    theme(legend.position="none") +
    labs(y="Price") 

# patchwork
(p2 | p1)
```

`r newthought("Theory E:")` Gas stations with direct highway access charge more.


`r newthought("Conclusion:")` The distribution of prices without direct highway access is more left-skewed than the price distribution with direct access. This is clear to see in the density plot on the right, but also when comparing the medians in the boxplot on the left.

\pagebreak
Also, we can make a map by geocoding the addresses. 

```{r gasmap, echo=TRUE, fig.cap="Map of Austin gas stations", fig.height=8, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE}
tryCatch({
  # Try: Run map
  atx_map <- ggmap(get_map("austin", zoom = 13), 
                   ylab="Latitude", 
                   xlab="Longitude") 

  m2 <- atx_map + 
    geom_point(data=gasprices, 
               mapping = aes(x=longitude, 
                             y=latitude, 
                            color=Brand)) + 
    scale_color_brewer(palette="Set1")
  
  m2 + labs(x="Longitude",
            y="Latitude") + 
    theme(text = element_text(family="Palatino"))
  # Catch: If replicating, need unique google API to work 
}, error = function(e) {
  "Error: Need Google map API to run map code for ggmap"
  }
)
```

\pagebreak 

# Data Visualization: A Bike Share Network

```{r bikeshare, warning=FALSE, include=FALSE}
bikeshare <- read_csv("../../ECO395M/data/bikeshare.csv")
```

## Plot 1 

```{r rents_v_hr, fig.height = 3, fig.width = 5, fig.cap="Bike rentals (total) versus hour of day (hr)"}
p1 <- bikeshare %>% 
  group_by(hr) %>% 
  mutate(avg_tot = mean(total)) %>% 
  ggplot( aes(x = hr) ) +
  geom_jitter(aes(y = total), color = palette_light()["blue"], alpha = 0.15)

p2 <- p1 + geom_line( aes(y = avg_tot), color = palette_light()["red"])

p3 <- p2 + 
  labs(x = "Hour of day (24hr)",
       y = "Average rentals") +
  theme_tufte()

p3
```

\pagebreak

## Plot 2

```{r rent_v_hour_wd, fig.fullwidth=TRUE, fig.height = 3, fig.width = 9, fig.cap="Bike rentals (total) versus hour of day (hr) by working day"}
workingday_labs <- c("Not working day", "Working day")
names(workingday_labs) <- c(0, 1)

p2 <- p1 + geom_smooth(aes(y = total), color = palette_light()["red"]) + 
  facet_grid(. ~ workingday,
                labeller = labeller(workingday = workingday_labs))  

p2 + 
  labs(x = "Hour of day (24hr)",
       y = "Average rentals") +
  theme_tufte()
```

\pagebreak 

## Plot 3

```{r 8am_weather, fig.cap="Average ridership during 8 AM hour by weather code and working day"}
g1 <- bikeshare %>% 
    filter(hr == 8) %>% 
    group_by(weathersit) %>% 
    mutate(avg_tot = mean(total)) %>% 
    ggplot() +
    geom_col(aes(x = weathersit, 
                 y = avg_tot, 
                 fill = factor(weathersit)))

weathersit_labs <- c("Clear", "Mist", "Light Snow")

g2 <- g1 + 
  scale_y_continuous(breaks = seq(0, 160000, by = 25000)) +
  labs(x = "Weather Situation",
       y = "Average rentals") +
  theme_tufte() +
  scale_fill_brewer(palette = "Set1", labels = weathersit_labs) +
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) 

g2 + facet_grid(. ~ workingday,
                labeller = labeller(workingday = workingday_labs))
```

\pagebreak

# Data visualization: Flights at ABIA

```{r flightsABIA, warning=FALSE, include=FALSE}
abia <- read_csv("../../ECO395M/data/ABIA.csv")

airport_codes <- readr::read_csv(url("https://ourairports.com/data/airports.csv")) %>% 
  select(iata_code, name, type, iso_country, 
         iso_region, municipality, latitude_deg, longitude_deg) %>% 
  rename(latitude = latitude_deg,
         longitude = longitude_deg)

# Join to get long/lat
abia <- abia %>% 
  left_join(airport_codes, by = c("Origin" = "iata_code")) %>% 
  rename(origin_name = name,
         origin_lat = latitude,
         origin_lon = longitude) %>%
  left_join(airport_codes, by = c("Dest" = "iata_code")) %>% 
  rename(dest_name = name,
         dest_lat = latitude,
         dest_lon = longitude) %>% 
  select(1:30, 35:37, 42:43)
```

```{r top5_tab, eval=FALSE, cache=TRUE, include=FALSE}
# Top five table
num_flights <- abia %>% 
  group_by(dest_name) %>% 
  tally() %>% 
  rename(Destination = dest_name,
         Count = n) %>% 
  filter(Count > 400)

color_me_a <- which(str_detect(num_flights$Destination, 
        "Dallas|Houston|Phoenix|Denver"))
color_me_b <- which(str_detect(num_flights$Destination, "Austin"))

knitr::kable(num_flights, 
             booktabs = TRUE,
             format = "latex",
             caption = "Number of flights in and out of Austin Bergstrom. Top 5 airports highlighted in pink",
             escape = FALSE
             ) %>% 
  kable_styling(latex_options = "hold_position") %>%
  row_spec(color_me_a, color = "black", background = "#fddaec") %>% 
  row_spec(color_me_b, color = "black", background = "#ccebc5")
```

Here, we visualize the flightpaths available in the dataset. We only observe domestic flights. Each vertex represents an airport, and each edge is a flight between Austin Bergstrom (green) and another airport.

```{r flightpath, echo=FALSE, cache=TRUE}

# US Map
if (require("maps")) {
  data(us.cities)
  capitals <- subset(us.cities, capital == 2)
  usmap <- ggplot(capitals, aes(long, lat)) +
    borders("state") +
    scale_size_area() +
    coord_quickmap()
}

# Get unique flight paths
abia_unique <- abia %>% distinct(origin_name, dest_name) %>% 
  left_join(abia, by= c("origin_name" = "origin_name", 
                        "dest_name" = "dest_name")) %>% 
  select(1:2, 32:35) %>% 
  unique()

# Get atx lon lat
atx_lon_lat <- abia_unique %>% 
  filter(str_detect(origin_name, "Austin")) %>% 
  select(origin_name, origin_lat, origin_lon) %>% 
  rename(atx_lat = origin_lat,
         atx_lon = origin_lon) %>% 
  distinct()

# Add flight paths
f1 <- usmap + 
  geom_curve(data = abia_unique, aes(x = origin_lon, y = origin_lat, 
                                   xend = dest_lon, yend = dest_lat), 
             col = "cadetblue", size = .5, curvature = .2) + 
  geom_point(data = abia_unique, aes(x = origin_lon, y = origin_lat), 
             col = "firebrick") +
  geom_point(data = atx_lon_lat, aes(x = atx_lon, y = atx_lat),
             col = "darkolivegreen", size = 2.5)

# theme
f1 + theme_map()
```

In an effort to narrow down the dataset, we want to find which carriers complete 90\% of the flights between Austin and any of these other destinations. We find that 10 out of 16 carriers complete 90\% of flights. For these purposes, 90\% is largely arbitrary, a more careful analysis would consider the impact of dropping these 6 carriers. For example, maybe a dropped carrier provides all of the flights to a particular region. We do not address those concerns here. 

```{r cumsum, echo=FALSE}
# find total number of flights by group and total
tot_flights <- abia %>% 
  group_by(UniqueCarrier) %>% 
  tally() %>% 
  as_tibble() %>% 
  select(2) %>% 
  sum()

# get top 90% of flights as df
abia90perc <- abia %>% 
  group_by(UniqueCarrier) %>% 
  tally() %>% 
  mutate(perc_flights = n / tot_flights) %>% 
  arrange(desc(perc_flights)) %>% 
  mutate(cum_sum = cumsum(perc_flights)) %>% 
  slice(1:10) 

# save list of top carriers
top10carriers <- abia90perc$UniqueCarrier

# print table of contribution by carrier
abia90perc %>% 
  kbl(
    booktabs = TRUE,
             format = "latex",
             caption = "Top carriers by number of flights",
             escape = TRUE
             ) %>% 
  kable_styling(latex_options = "hold_position") 

# save a df filtering out bottom carriers
abia_topcarriers <- abia %>% filter(UniqueCarrier %in% top10carriers)
```

A carrier delay is where the cause of the cancellation or delay was due to circumstances within the airline's control (e.g. maintenance or crew problems, aircraft cleaning, baggage loading, fueling, etc.). What is the probability of carrier delay by an airline?

```{r mindelay, echo=FALSE, cache=TRUE, fig.cap="Probability of delay by carrier", message=FALSE, warning=FALSE, fig.width=5, fig.height=3}

pr1 <- abia_topcarriers %>% 
  group_by(UniqueCarrier) %>% 
  add_tally() %>% 
  rename(tot_flight = n) %>% 
  mutate(mean_carr_delay = round(mean(CarrierDelay, na.rm = TRUE), 2),
         CarrierDelay_binary = if_else(CarrierDelay > 0, 1, 0)) %>% 
  select(9, 36:37, 38) %>% 
  mutate(num_carr_delay = sum(CarrierDelay_binary, na.rm = TRUE)) %>% 
  select(-4) %>% 
  distinct(UniqueCarrier, .keep_all = TRUE) %>% 
  mutate(prob_carr_delay = round((num_carr_delay / tot_flight)*100, 3))

pr2 <- pr1 %>% 
  ggplot() +
  geom_col(aes(reorder(x = UniqueCarrier, -prob_carr_delay), 
               y = prob_carr_delay), fill = palette_dark()["blue"], 
           alpha = 0.75, color = "black") +
  scale_y_continuous(labels=function(x) paste0(x, "%")) 

pr3 <- pr2 +
  labs(x = "Carrier", y = "Probability of Carrier Delay") + 
  theme_tufte()

pr3
```

Here, we show the probability of carrier delay by day of week. Mondays and Fridays have the highest likelihood of a carrier delay.

```{r probdelay_day, echo=FALSE, fig.cap="Probability of carrier delay by day of week", fig.height=3, fig.width=5, cache=TRUE}
pdd1 <- abia_topcarriers %>% 
  group_by(UniqueCarrier, DayOfWeek) %>% 
  add_tally() %>% 
  rename(tot_flight = n) %>% 
  mutate(mean_carr_delay = round(mean(CarrierDelay, na.rm = TRUE), 2),
         CarrierDelay_binary = if_else(CarrierDelay > 0, 1, 0)) %>% 
  select(4, 9, 36:37, 38) %>% 
  mutate(num_carr_delay = sum(CarrierDelay_binary, na.rm = TRUE)) %>% 
  select(-5) %>% 
  distinct(UniqueCarrier, DayOfWeek, .keep_all = TRUE) %>% 
  mutate(prob_carr_delay = round((num_carr_delay / tot_flight)*100, 3)) %>% 
  ungroup() 

DayOfWeek_lab <- c("1" = "Mon", "2" = "Tues", "3" = "Wed", "4" = "Thurs", 
                   "5" = "Fri", "6" = "Sat", "7" = "Sun")

pdd2 <- pdd1 %>% 
  group_by(DayOfWeek) %>% 
  mutate(prob_carr_delay = sum(prob_carr_delay)) %>% 
  distinct(prob_carr_delay) %>% 
  ggplot() +
  geom_col(aes(x = factor(DayOfWeek), y = prob_carr_delay), 
           fill = palette_dark()["blue"], 
           alpha = 0.75, color = "black") +
  scale_y_continuous(labels=function(x) paste0(x / 10, "%")) +
  scale_x_discrete(labels = DayOfWeek_lab)

pdd3 <- pdd2 +
  labs(x = "Day of Week", y = "Probability of Carrier Delay") + 
  theme_tufte() 

pdd3
```

It seems that flying on Mondays and Fridays is associated with the highest probability of encountering a carrier delay. We could change the `delay` variable and repeat this analysis for all other delays if we wish.

\pagebreak

# K-Nearest Neighbors

```{r sclass, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
sclass <- read_csv("../../ECO395M/data/sclass.csv") %>% 
  filter(trim %in% c("350", "65 AMG"))

# Filter data
sclass350 <- sclass %>% filter(trim == "350")
sclass65 <- sclass %>% filter(trim == "65 AMG")

# Table
sclass %>% 
  group_by(trim) %>% 
  tally() %>% 
  kbl(booktabs = TRUE,
             format = "latex",
             caption = "Count by trim",
             escape = FALSE
             ) %>% 
  kable_styling(latex_options = "hold_position")
```

```{r eval=FALSE, cache=TRUE, message=FALSE, warning=FALSE, include=FALSE}
sclass %>% 
  ggplot() + 
  geom_point(aes(x = mileage, y = price, color = trim))
```

```{r traintest, cache=TRUE, echo=FALSE}
set.seed(350)

# split into train/test
sclass350_split <- initial_split(sclass350, prop = 0.6, strata = price)
sclass350_train <- training(sclass350_split)
sclass350_test <- testing(sclass350_split)

sclass65_split <- initial_split(sclass65, prop = 0.6, strata = price)
sclass65_train <- training(sclass65_split)
sclass65_test <- testing(sclass65_split)
```

## sclass 350

```{r modspec350, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
# 350
sclass350_recipe <- recipe(price ~ mileage, data = sclass350_train) %>%
  step_scale(all_predictors()) %>%
  step_center(all_predictors())

sclass350_spec <- nearest_neighbor(weight_func = "rectangular", neighbors = tune()) %>%
  set_engine("kknn") %>%
  set_mode("regression")

sclass350_vfold <- vfold_cv(sclass350_train, v = 2, strata = price)

sclass350_wkflw <- workflow() %>%
  add_recipe(sclass350_recipe) %>%
  add_model(sclass350_spec)
```

```{r res350, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
gridvals <- tibble(neighbors = seq(1, 124))

sclass350_results <- sclass350_wkflw %>%
  tune_grid(resamples = sclass350_vfold, grid = gridvals) %>%
  collect_metrics()

# show all the results
# sclass350_results
```


```{r rmse_x_k_350, cache=TRUE, echo=FALSE, fig.cap="RMSE versus k", fig.height=3, fig.width=5, warning=FALSE}
kp1 <- sclass350_results %>% 
  filter(.metric == "rmse") %>% 
  select(neighbors, mean) %>% 
  mutate(highlight = if_else(mean == min(mean), "h", "l")) %>% 
  ggplot() + 
  geom_point(aes(x = neighbors, y = mean, color = highlight)) + 
  scale_color_manual(values = c("tomato", "grey")) +
  xlim(0, 75) + 
  labs(x = "Neighbors, K", y = "RMSE") 

kp2 <- kp1 + 
  theme_tufte() + 
  theme(legend.position = "none")

kp2
```


```{r min_rmse_350, cache=TRUE, echo=FALSE}
# show only the row of minimum RMSPE
sclass350_min <- sclass350_results %>%
  filter(.metric == "rmse") %>%
  filter(mean == min(mean)) 

sclass350_min %>%  
    kbl(booktabs = TRUE,
             format = "latex",
             caption = "Minimum RMSE",
             escape = TRUE
             ) %>% 
  kable_styling(latex_options = "hold_position")

print(paste("The smallest rmse occurs at K =", as.numeric(sclass350_min[,1])))
```

```{r evaltest_350, cache=TRUE, include=FALSE}
set.seed(350)
kmin <- sclass350_min %>% pull(neighbors)
sclass350_spec <- nearest_neighbor(weight_func = "rectangular", neighbors = kmin) %>%
  set_engine("kknn") %>%
  set_mode("regression")

sclass350_fit <- workflow() %>%
  add_recipe(sclass350_recipe) %>%
  add_model(sclass350_spec) %>%
  fit(data = sclass350_train)

sclass350_summary <- sclass350_fit %>%
  predict(sclass350_test) %>%
  bind_cols(sclass350_test) %>%
  metrics(truth = price, estimate = .pred)

sclass350_summary
```

```{r pred_350, cache=TRUE, echo=FALSE, fig.height = 3, fig.width = 5, fig.cap="Overlaid fitted model for optimal value of K, where red is the 65 AMG trim and blue is the 350 trim."}
set.seed(350)
sclass350_preds <- sclass350_fit %>%
  predict(sclass350_train) %>%
  bind_cols(sclass350_train)

plot_final <- ggplot(sclass350_preds, aes(x = mileage, y = price)) +
  geom_point(alpha = 0.4) +
  xlab("Mileage") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_line(data = sclass350_preds, aes(x = mileage, y = .pred), 
            color = "tomato") +
  ggtitle(paste0("K = ", kmin)) 

plot_final + theme_tufte()
```


\pagebreak

## sclass 65 AMG

```{r modspec65, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
# 65 AMG
sclass65_recipe <- recipe(price ~ mileage, data = sclass65_train) %>%
  step_scale(all_predictors()) %>%
  step_center(all_predictors())

sclass65_spec <- nearest_neighbor(weight_func = "rectangular", neighbors = tune()) %>%
  set_engine("kknn") %>%
  set_mode("regression")

sclass65_vfold <- vfold_cv(sclass65_train, v = 2, strata = price)

sclass65_wkflw <- workflow() %>%
  add_recipe(sclass65_recipe) %>%
  add_model(sclass65_spec)
```

```{r res65, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
gridvals <- tibble(neighbors = seq(1, 50))

sclass65_results <- sclass65_wkflw %>%
  tune_grid(resamples = sclass65_vfold, grid = gridvals) %>%
  collect_metrics()

# show all the results
# sclass350_results
```


```{r rmse_x_k_65, cache=TRUE, echo=FALSE, fig.height = 3, fig.width = 5, fig.cap="RMSE versus K"}
kp1 <- sclass65_results %>% 
  filter(.metric == "rmse") %>% 
  select(neighbors, mean) %>% 
  mutate(highlight = if_else(mean == min(mean), "h", "l")) %>% 
  ggplot() + 
  geom_point(aes(x = neighbors, y = mean, color = highlight)) + 
  scale_color_manual(values = c("tomato", "grey")) +
  xlim(0, 75) + 
  labs(x = "Neighbors, K", y = "RMSE") 

kp2 <- kp1 + 
  theme_tufte() + 
  theme(legend.position = "none")

kp2
```


```{r min_rmse_65, cache=TRUE, echo=FALSE}
# show only the row of minimum RMSPE
sclass65_min <- sclass65_results %>%
  filter(.metric == "rmse") %>%
  filter(mean == min(mean)) 

sclass65_min %>%  
    kbl(booktabs = TRUE,
             format = "latex",
             caption = "Minimum RMSE",
             escape = TRUE
             ) %>% 
  kable_styling(latex_options = "hold_position")

print(paste("The smallest rmse occurs at K =", as.numeric(sclass65_min[,1])))
```

```{r evaltest_65, cache=TRUE, include=FALSE}
set.seed(350)
kmin <- sclass65_min %>% pull(neighbors)
sclass65_spec <- nearest_neighbor(weight_func = "rectangular", neighbors = kmin) %>%
  set_engine("kknn") %>%
  set_mode("regression")

sclass65_fit <- workflow() %>%
  add_recipe(sclass65_recipe) %>%
  add_model(sclass65_spec) %>%
  fit(data = sclass65_train)

sclass65_summary <- sclass65_fit %>%
  predict(sclass65_test) %>%
  bind_cols(sclass65_test) %>%
  metrics(truth = price, estimate = .pred)

sclass65_summary
```

```{r pred_65, cache=TRUE,echo=FALSE, fig.height = 3, fig.width = 5, fig.cap="Fitted model for optimal value of K"}
set.seed(350)
sclass65_preds <- sclass65_fit %>%
  predict(sclass65_train) %>%
  bind_cols(sclass65_train)

plot_final <- ggplot(sclass65_preds, aes(x = mileage, y = price)) +
  geom_point(alpha = 0.4) +
  xlab("Mileage") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_line(data = sclass65_preds, aes(x = mileage, y = .pred), 
            color = "tomato") +
  ggtitle(paste0("K = ", kmin)) 

plot_final + theme_tufte()
```

It is also interesting to look at both `mileage` versus `price` graphs together.

```{r pred_65_350, cache=TRUE,echo=FALSE, fig.height = 3, fig.width = 5, fig.cap="Fitted model for optimal value of K"}
plot_joint <- ggplot() +
  geom_point(data = sclass65_preds, aes(x = mileage, y = price), color = "tomato", alpha = 0.4) +
  geom_point(data = sclass350_preds, aes(x = mileage, y = price), color = "dodgerblue", alpha = 0.4) +
  xlab("Mileage") +
  ylab("Price (USD)") +
  scale_y_continuous(labels = dollar_format()) +
  geom_line(data = sclass65_preds, aes(x = mileage, y = .pred, color = "65 AMG")) +
  geom_line(data = sclass350_preds, aes(x = mileage, y = .pred, color = "350")) + 
  scale_color_manual(values = c("blue", "red"))

plot_joint + 
  theme_tufte() + 
  theme(legend.title = element_blank())
```


```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```


