---
title: "394D PS1"
subtitle: "Week 1"
author: "Scott Cohn"
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_html: default
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
library(tidyverse)
library(kableExtra)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

# 1. Used Car Dealer

A used car dealer has 30 cars and 10 of them are lemons. The total number of cars is 30, $N = 30$. The marginal probability of getting a lemon is $P(L) = \frac{\text{\# of Lemons}}{N} =\frac{10}{30} = \frac{1}{3}$. Similarly, the marginal probability of getting a non-Lemon $P(\neg L) = \frac{20}{30} = \frac{2}{3}$. We are interested in the probability of getting _at least_ one lemon. Let $X$ denote the number of lemons bought so we want $P(X \geq 1)$.

We want to purchase all three cars simultaneously. We take the total number of chances of 1, 2, or 3 lemons divided by the total number of buying combinations.

\[
  \frac{\binom{20}{2} \binom{10}{1} + \binom{20}{1}\binom{10}{2} + \binom{20}{0}\binom{10}{3}}{\binom{30}{3}} = \frac{2800}{4060} = 0.69 = 69 \%
\]

So there is a 69\% chance of buying at least one lemon.


# 2. Two Dice

We throw two dice. 

## What is probability that the sum of the two numbers is odd? 

A single die has values 1-6, so the sum of any two die is in the closed interval $[2, 12]$. However, since each die has 6 sides, there are 36 possible combinations to achieve one of those sums. Thus, $N = 36$. Of those 36, half are odd --- $n_{\text{odd}} = 18$. 

It follows that the probability that the sum of the two numbers is odd is:

\[
  P(\text{sum is odd}) = \frac{n_{\text{odd}}}{N} = \frac{18}{36} = \frac{1}{2}.
\]

## What is the probability that the sum of the two numbers is less than 7?

There are 15 possible combinations of rolls, denoted set $\Omega$, where the sum is strictly less than 7:

\[
  \Omega = \{(1, 1), (1, 2), (2, 1), (1, 3), (3, 1), (1, 4), (4, 1), (1, 5), (5, 1), (2,2), (2, 3), (3, 2), (3, 3), (2, 4), (4, 2)\}
\]

Therefore, 

\[
  P(\text{sum} < 7) = \frac{\text{length}(\Omega)}{N} = \frac{15}{36} = \frac{5}{12} \approx 42\%
\]

## What is the probability that the sum of the two numbers is less than 7 given that it is odd? 

The above is shown mathematically as $P(\text{sum} < 7 | \text{sum is odd})$. Let $A$ denote the statement "sum $< 7$" and $B$ denote the statement "sum is odd". Then,

\[P(A|B) = \frac{P(A \cap B)}{P(B)}\]

There are 18 cases where the sum is odd: 

\begin{align*}
\Omega &= \{(1, 2), (2,1), (1,4), (4,1), (1,6), (6,1), (2,3), (3,2), \\
  & (2,5), (5,2), (3,4), (4,3), (3,6), (6,3), (4,5), (5,4), (5,6), (6,5)\}
\end{align*}

We know from above that $P(B) = \frac{18}{36} = \frac{1}{2}$. It remains to find $P(A \cap B)$. 

We want the intersection of statement $A$ and statement $B$. Thus, $P(A \cap B)$ is equal to all of the sums in set $\Omega$ where the sum is less than 7 divided by 36 total combinations. 

\[A \cap B = \{\omega \in \Omega : \omega \text{ is odd sum and } \omega \text{ sum is} < 7\}\]

There are 6 such combinations of odds with a sum less than 7:

\[\Omega_{\omega < 7} = \{(1, 2), (2,1), (1,4), (4,1), (2,3), (3,2)\}\] 

So, $P(A \cap B) = \frac{6}{36} = \frac{1}{6}$. Putting it all together, we see that 

\[P(A|B) = \frac{P(A \cap B)}{P(B)} = \frac{1/6}{1/2} = \frac{1}{3} \approx 33\%\]

## Are these two events independent?

If the events are independent, then $P(A \cup B) = P(A) \times P(B)$. We also know that $P(A \cup B) = P(A) + P(B) - P(A \cap B)$. If these evaluate to be the same value, then events $A$ and $B$ are independent.

We know the following:

- $P(A) = \frac{5}{12}$
- $P(B) = \frac{1}{2}$
- $P(A \cap B) = \frac{1}{6}$

Thus,

\begin{align*}
  P(A \cup B) &= P(A) \times P(B) \\
  &= \frac{5}{12} \times \frac{1}{2} \\
  &= \frac{5}{24}
\end{align*}

\begin{align*}
  P(A \cup B) &= P(A) + P(B) - P(A \cap B) \\
  &= \frac{5}{12} + \frac{1}{2} - \frac{1}{6} \\
  &= \frac{3}{4}
\end{align*}

It clearly follows that $\frac{5}{24} \neq \frac{3}{4}$. Hence, the two events are _not_ independent. Before doing any work, we can also observe that because their intersection is non-zero, the events are independent.

# 3. Random Clicker and Truthful Clicker

There are two types of users that access the website: RC and TC. RC's make up 30\% of users. Then, TC = 1 - RC = 70\%.

The marginal probabilities for user-type are:

- $P(TC) = 0.7$
- $P(RC) = 0.3$

It is given that RC-types will click Yes/No equally. Then the conditional probabilities for RC are:

- $P(Y|RC) = 0.5$
- $P(N|RC) = 0.5$

Using the marginal and conditional probabilities for RC-types, we compute the joint probabilities.

- $P(Y \cap RC) = P(RC) P(Y | RC) = 0.15$
- $P(N \cap RC) = P(RC) P(N | RC) = 0.15$

\begin{marginfigure}
  \centering
  \includegraphics{Figures/394_ps1_q3.eps}
  \caption{Tree for TC and RC}
  \label{fig:tree}
\end{marginfigure}

These probabilities are visualized in figure \ref{fig:tree}. It remains to find the conditional and joint probabilities for the TC-type. After a trial period, we know 65\% said Yes and 35\% said No. It follows that:

\begin{align}
  P(Y) &= P(Y \cap TC) + P(Y \cap RC) \\
  0.65 &= P(Y \cap TC) + 0.15 & \text{Plugging in.} \\
  \therefore P(Y \cap TC) &= 0.5
\end{align}

Similarly,

\begin{align}
  P(N) &= P(N \cap TC) + P(N \cap RC) \\
  0.35 &= P(N \cap TC) + 0.15 & \text{Plugging in.} \\
  \therefore P(N \cap TC) &= 0.2
\end{align}

Using the joint probability $P(Y \cap TC)$, we can compute $P(Y | TC)$ using the following formula:

\[
  P(Y | TC) = \frac{P(Y \cap TC)}{P(TC)} = \frac{0.5}{0.7} \approx 0.714
\]

We interpret $P(Y | TC)$ as the probability that a person clicked yes _given_ that they are a truthful clicker. Equivalently, this is the fraction of people who are truthful clickers that answered yes. About 71\% of truthful clickers (TC) clicked yes.

# 4. Medical Test

`r margin_note("Well this is a topical question ...")`

The marginal probabilities for having the disease are:^[taken from bullet 3]

- $P(D) = 0.000025$
- $P(\neg D) = 1 - P(D) = 0.999975$

From the sensitivity, we know the conditional probability of testing positive given that an individual has the disease is $P(+|D) = 0.993$. We can then calculate the conditional probability of testing negative given that an individual has the disease: $P(-|D) = 1 - P(+|D) = 0.007$. Thus, we have:

- $P(+|D) = 0.993$
- $P(-|D) = 0.007$

Similarly, we use the specificity to get $P(- | \neg D) = 0.9999$ and $P(+ | \neg D) = 1 - P(- | \neg D) = 0.0001$.

- $P(- | \neg D) = 0.9999$
- $P(+ | \neg D) = 0.0001$

Multiplying the marginal probabilities by the conditional probabilities yields the following joint probabilities:
`r margin_note("We can check these are correct by making sure they sum to 1")`

- $P(+ \cap D) = 0.00002482$
- $P(- \cap D) = 0.00000018$
- $P(+ \cap \neg D) = 0.0001$
- $P(- \cap \neg D) = 0.999875$

Now supposing someone tests positive, we want the probability that they have the disease: $P(D | +)$. To find this probability, we use Bayes:

\[
  P(D | +) = \frac{P(D)P(+|D)}{P(+)}
\]

We don't have the marginal probability of being positive, so we need to expand the denominator as follows:

\[
  P(D | +) = \frac{P(D)P(+|D)}{P(D)P(+|D) + P(\neg D)P(+| \neg D)}
\]

Plugging in the values from above yields the following:
\[
  P(D | +) = \frac{(0.000025)(0.993)}{(0.000025)(0.993) + (0.999975)(0.0001)} = 0.19888 \approx 20 \%
\]

This result indicates that if an individual were to test positive then it is a true positive only 20 \% of the time. This suggests a false positive rate that is quite large. If this is the case, an epidemiologist may ask "What if we're immune and don't know it?". To limit the rate of false positives, researchers would likely have to sample in regions where they know the disease is more prevalent, thus forcing a sample bias. Or, through very large sampling, these false positives could be factored out.

The primary concern is poor information. With high false positives, researchers are unable to accurately gauge progress in the fight against disease spread. Furthermore, patients may be exposed to treatments that they may not need. Depending on the treatment, this could be harmful. Alternatively, admittance to a hospital under the auspice of a being positive may in fact expose them to the disease when they were in fact negative in the first place.

# 5. One Match to Go^[Data from \textcolor{magenta}{\url{https://www.soccerstats.com/latest.asp?league=england}}]

Spiegelhalter and Ng's goal was to accurately predict soccer match results based on deviations from the norm for attacking and defense statistics, while controlling for home and away effects. They define "attack strength" as the ratio of goals scored by a team over the average number of goals scored by a team. Similarly, "defense weakness" is the ratio of goals that the team has let (or been scored against) over the average across the league. They also find the average number of goals scored by a home team and an away team to create a baseline. 

Then, they ask "how many goals do we expect a team to score?" The expected number of goals comes from weighting the average home or away benchmark by the respective team's "attack strength" and the opposition team's "defense weakness". The expectation allows the authors to use a Poisson distribution to find the probability that a team will reach a certain score in a given match-up. Furthermore, this allows the analyst to determine the probability of a given score by multiplying the each team's score probability together. Spiegelhalter and Ng use this method to compute the first, second, and third most likely result of match-ups based on scoring history. Teasing these values further, Spiegelhalter and Ng are able to construct a table of the percent probability of a home win, away win, or draw. Ex-post, they check the accuracy of their predictions with a Brier penalty, which is typically used in weather forecasting.

```{r gf_ga, echo = FALSE}

# df <- tibble(
#   team = c("Liverpool", "Tottenham"),
#   points = c(93, 52), # Earned by winning matches
#   # Liverpool is at the top, Tott upper mid
#   goals_for = c(76, 54),
#   goals_against = c(27, 45)
# )
df <- tibble(
  team = c("Liverpool", "Manchester C.", "Chelsea", "Leicester City", "Manchester Utd", "Wolverhampton", "Sheffield Utd", "Tottenham", "Arsenal", "Burnley", "Everton", "Southampton", "Newcastle Utd", "Crystal Palace", "Brighton", "West Ham Utd", "Watford", "Bournemouth", "Aston Villa", "Norwich City"),
  points = c(93, 72, 60, 59, 59, 55, 54, 52, 50, 50, 45, 45, 43, 42, 36, 34, 34, 31, 30, 21), # Earned by winning matches
  goals_for = c(76, 91, 63, 65, 61, 48, 38, 54, 51, 39, 41, 45, 36, 30, 36, 44, 33, 36, 38, 26),
  goals_against = c(27, 34, 49, 36, 35, 37, 33, 45, 44, 47, 52, 58, 52, 45, 52, 59, 54, 60, 65, 67)
)
# Find average number of goals scored by a team
gf_league <- c(91, 76, 65, 63, 61, 54, 51, 48, 45, 44,41,39, 38, 38, 36, 36, 36, 33, 30, 26)
gf_avg <- mean(gf_league)

df <- df %>% mutate(
  attack_strength = goals_for / gf_avg,
  defense_weakness = goals_against / gf_avg
    )

# reorder cols by index
df <- df[c(1, 2, 3, 5, 4, 6)]

# Pulled from table on site
home_goals_avg <- 1.52
away_goals_avg <- 1.2
```


```{r table1, echo = FALSE}
col_names <- c("Team", "Points", linebreak("Goals\nfor"), linebreak("Attack\nstrength"), linebreak("Goals\nagainst"), linebreak("Defense\nweakness"))

color_me_a <- which(df$team %in% c("Liverpool", "Tottenham"))
color_me_b <- which(df$team %in% c("Manchester C.", "Manchester Utd"))

knitr::kable(df, col.names = col_names, 
             digits = 2,
             booktabs = TRUE,
             format = "latex",
             caption = "State of the Premier League as of \\textcolor{RoyalBlue}{\\textbf{July 13, 2020}}",
             escape = FALSE
             ) %>% 
  kable_styling(latex_options = "hold_position") %>%
  row_spec(color_me_a, color = "black", background = "#fddaec") %>% 
  row_spec(color_me_b, color = "black", background = "#ccebc5")
  #kable_styling(bootstrap_options = "striped")
```


Home teams scored on average 1.52 goals per match, while away teams scored 1.2 goals per match.^[Given in online table, not calculated] 

First, we wish the predict the result of Liverpool (home) _versus_ Tottenham (away). If Liverpool were average, we would expect them to score 1.52 goals. Liverpool is not average and scores 160\% of the average number of goals.^[Liverpool's "attack strength" is 1.60] Multiplying $1.52 \times 1.60 = 2.43$. We expect them to score 2.43 goals against an average team. However, we must account for the defense of Tottenham, the away team. Tottennam's "defense weakness" is 0.95. That is, they concede 95\% of the average goals scored. Amending the previous calculation by factoring in the defense, $1.52 \times 1.60 \times 0.95 = 2.31$ expected goals by Liverpool.

We apply a similar approach to the Tottenham squad: Amending the baseline of 1.2 yields $1.2 \times 1.14 \times 0.57 = 0.78$ expected goals in a match against Liverpool.

```{r poisson, echo = FALSE}
# taken from above -- expected goals in a game against each other
liverpool_lambda = 2.43
tottenham_lambda = 0.78

# vector of scores
x <- seq(0, 5, by = 1)

# create tibble
df_poisson <- tibble(
  team = c("Liverpool", "Tottenham"),
  exp_goal = c(liverpool_lambda, tottenham_lambda)
)

# add probabilities
df_poisson <- df_poisson %>% 
  mutate("0" = round(dpois(0, exp_goal), 2)*100,
         "1" = round(dpois(1, exp_goal), 2)*100,
         "2" = round(dpois(2, exp_goal), 2)*100,
         "3" = round(dpois(3, exp_goal), 2)*100,
         "4" = round(dpois(4, exp_goal), 2)*100,
         "5" = round(dpois(5, exp_goal), 2)*100)

# create table
col_names <- c("Team", "Expected goals", "0", "1", "2", "3", "4", "5")

knitr::kable(df_poisson, format = "latex",
             col.names = col_names, 
             digits = 2,
             booktabs = TRUE,
             caption = "Expected number of goals, and percentage chance of getting a particular score for the two teams, assuming a Poisson distribution"
             ) %>% 
  kable_styling(latex_options = "hold_position")
```

Next, we want to replicate the above analysis for Manchester United (home) _versus_ Manchester City (away). Again, the home team --- if average --- will score 1.52 goals a game. Manchester United has an "attack strength" of 1.28 and Manchester City has a "defense weakness" of 0.72. Thus, we expect Manchester United to score $1.52 \times 1.28 \times 0.72 = 1.4$ goals per game. Conversely, the average away team will score 1.2 goals per game, so adjusting by Manchester United's "defense weakness" (0.74) and Manchester City's "attack strength" (1.91) yield an expected $1.2 \times 1.91 \times 0.74 = 1.7$ goals per game.

```{r poisson2, echo = FALSE}
# taken from above -- expected goals in a game against each other
manc_lambda = 1.7
manu_lambda = 1.4

# vector of scores
x <- seq(0, 5, by = 1)

# create tibble
df_poisson <- tibble(
  team = c("Manchester Utd", "Manchester C."),
  exp_goal = c(manu_lambda, manc_lambda)
)

# add probabilities
df_poisson <- df_poisson %>% 
  mutate("0" = round(dpois(0, exp_goal), 2)*100,
         "1" = round(dpois(1, exp_goal), 2)*100,
         "2" = round(dpois(2, exp_goal), 2)*100,
         "3" = round(dpois(3, exp_goal), 2)*100,
         "4" = round(dpois(4, exp_goal), 2)*100,
         "5" = round(dpois(5, exp_goal), 2)*100)

# create table
col_names <- c("Team", "Expected goals", "0", "1", "2", "3", "4", "5")

knitr::kable(df_poisson, format = "latex",
             col.names = col_names, 
             digits = 2,
             booktabs = TRUE,
             caption = "Expected number of goals, and percentage chance of getting a particular score for the two teams, assuming a Poisson distribution"
             ) %>% 
  kable_styling(latex_options = "hold_position")
```

`r margin_note("See Table 4 for assessed probabilities for the 2 matches.")`

```{r win_loss_draw, message=FALSE, warning=FALSE}
liverpool <- c(0.09, 0.21, 0.26, 0.21, 0.13, 0.06)
tottenham <- c(0.46, 0.36, 0.14, 0.04, 0.01, 0.00)

manu <- c(0.25, 0.35, 0.24, 0.11, 0.04, 0.01)
manc <- c(0.18, 0.31, 0.26, 0.15, 0.06, 0.02)

# win_loss_draw matrix
# cols are liverpool rows are tott
wld_lt <- matrix(rep(NA,36),ncol=6)
# cols are manu rows are manc
wld_mm <- matrix(rep(NA,36),ncol=6)

for(i in 1:6)
  for(j in 1:i)
    wld_lt[i,j] <- liverpool[i]*tottenham[j]

for(i in 1:6)
  for(j in 1:i)
    wld_mm[i,j] <- manu[i]*manc[j]

# !! The values in the rows/cols are indexes, not scores !!
# e.g. (1, 1) is actually the P() for score (0, 0)
print(wld_lt, na.print = "")
print(wld_mm, na.print = "")

# replace na's with 0's to row-sum
# !! note values aren't actually 0 !!
wld_lt[is.na(wld_lt)] = 0
wld_mm[is.na(wld_mm)] = 0

# trace for probability of draws
prob_of_draw_lt <- sum(diag(wld_lt)) 
prob_of_draw_mm <- sum(diag(wld_mm)) 

# find sum of lower triangle, not including diagonal
prob_of_home_win_lt <- sum(rowSums(wld_lt * lower.tri(wld_lt, diag=FALSE)))
prob_of_home_win_mm <- sum(rowSums(wld_mm * lower.tri(wld_mm, diag=FALSE)))

# prob of home_loss = 1 - prob_win - prob_draw
prob_of_home_loss_lt <- 1 - prob_of_home_win_lt - prob_of_draw_lt
prob_of_home_loss_mm <- 1 - prob_of_home_win_mm - prob_of_draw_mm
```

```{r wld_table, echo = FALSE}
wld_df <- tibble(
  home = c("Liverpool", "Manchester Utd"),
  away = c("Tottenham", "Manchester C."),
  home_win = c(prob_of_home_win_lt, prob_of_home_win_mm),
  draw = c(prob_of_draw_lt, prob_of_draw_mm),
  away_win = c(prob_of_home_loss_lt, prob_of_home_loss_mm)
)

# create table
col_names_wld <- c("Home", "Away", "Home win", "Draw", "Away win")

knitr::kable(wld_df, format = "latex",
             col.names = col_names_wld, 
             digits = 2,
             booktabs = TRUE,
             caption = "The assessed probabilities for the 2 matches."
             ) %>% 
  kable_styling(latex_options = "hold_position")
```

```{r eval=FALSE, message=FALSE, warning=FALSE, echo = FALSE}
file.edit(
  tufte:::template_resources(
    'tufte_html', '..', 'skeleton', 'skeleton.Rmd'
  )
)
```

```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```
